Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB

### 2018-07-10 paper/paper.tex
ready to make pdf

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_mechanism}
	\caption{Modeling intracellular mechanisms. (box A) A single-compartment neuron model with 8 channel types. With these parameters, this model neuron is a tonic spiker. Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (top panel). (box B)  An intracellular buffering of Calcium and the effect of Calcium influx through voltage gated Calcium channels (VGCCs) is modeled using a simple differential equation. Code snippet shows how this mechanism can be added to the neuron model. The cell now bursts periodically, with synchronized oscillations in intracellular Calcium. (box C) Schematic of Calcium-dependent integral control homeostasis. In this feedback system, the rates of production of mRNA depends on the Calcium levels in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance. These integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances. On integrating the model, intracellular calcium levels rise and approach the target (red dashed line). This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism. The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}
	\label{fig:figuremechanism}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad  arbitrium$. The initial state (orange disc) of the single compartment model in the previous example is saved using the \texttt{snapshot} method. This method saves all parameters and dynamic variables of the model in a named state. The first column shows the profile of conductances and the voltage dynamics of the model at this point in time. The maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron (purple plots). After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots). This configuration is now saved in a state called $final$ and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange). Another parameter is changed (the Calcium target), and the model is integrated to reach a new state (blue disc) where the voltage dynamics are different from the $initial$ state. In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
	\label{fig:figuresnapshot}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
	\caption{Simulating a network of conductance-based model neurons. The diagram shows a three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004). Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances. Synapses are of two types (Cholinergic (dashed) or Glutamatergic (solid lines)) and have different kinetics. The code snippet shows how neurons can be created, wired together using synapses, and the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse. (A-C) Simulated voltage trace of a model network for the three compartments obtained from this simulation. (D) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
	\label{fig:figurenetwork}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_manipulate}
	\caption{Manipulating neuron parameters in real time. Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Figure 1 is being manipulated using the code snippet shown here. A GUI with sliders for each parameter being manipualted is created, and linked to an atbitrary number of visualization functions. In this example, the visualization functions are the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-vs.-injected current curve for this neuron (B). Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how that parameter controls the dynamical behavior of the model. A screen recording of this model being manipulated in real time is included in Supplementary Material XX. }
	\label{fig:figuremanipulate}
\end{figure}

>>>>

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_mechanism}
	\caption{Modeling intracellular mechanisms. (box A) A single-compartment neuron model with 8 channel types. With these parameters, this model neuron is a tonic spiker. Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (top panel). (box B)  An intracellular buffering of Calcium and the effect of Calcium influx through voltage gated Calcium channels (VGCCs) is modeled using a simple differential equation. Code snippet shows how this mechanism can be added to the neuron model. The cell now bursts periodically, with synchronized oscillations in intracellular Calcium. (box C) Schematic of Calcium-dependent integral control homeostasis (cite O'Leary et al.). In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance. These integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances. On integrating the model, intracellular calcium levels rise and approach the target (red dashed line). This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism. The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}
	\label{fig:figuremechanism}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad$   $arbitrium$. The initial state (orange disc) of the single compartment model in the previous example is saved using the \texttt{snapshot} method. This method saves all parameters and dynamic variables of the model in a named state. The first column shows the profile of conductances and the voltage dynamics of the model at this point in time. The maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron (purple plots). After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots). This configuration is now saved in a state called $final$ and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange). Another parameter is changed (the Calcium target), and the model is integrated to reach a new state (blue disc) where the voltage dynamics are different from the $initial$ state. In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
	\label{fig:figuresnapshot}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
	\caption{Simulating a network of conductance-based model neurons. The diagram shows a three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004). Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances. Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics. The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse. (A-C) Simulated voltage trace of a model network for the three compartments obtained from this simulation. (D) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
	\label{fig:figurenetwork}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_manipulate}
	\caption{Manipulating neuron parameters in real time. Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Figure 1 is being manipulated using the code snippet shown here. The screenshot shows a GUI with sliders for every parameter of interest that is created by the \texttt{manipulate} method, and linked to an arbitrary number of visualization functions. In this example, two visualization functions are used: the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-$vs$.-injected current curve for this neuron (B). Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model. A screen recording of this model being manipulated in real time is included in Supplementary Material XX. }
	\label{fig:figuremanipulate}
\end{figure}
